Model_additionalAlts

### 2015-12-01 models/Model_additionalAlts.Rmd
modiefied gradient descent tracking and plotting

Focus on single scalar fn()
```{r}
# focus_alts
# ---------
# target    --> specific scalar family we're interested in (i.e. "some_all")
# combos    --> collection of possible 'mid' l0 distribution
# normalize --> normaize pre RSA?
# alpha     --> set alpha level
focus_alts = function(target, combos, normalize=F, alpha=1) {
  # Default normalish looking distr
  if (is.null(combos)) {
    combos = data.frame(c(0, 0, 1, 0, 0),
                        c(0, 0.05, 0.9, 0.05, 0),
                        c(0, 0.1, 0.8, 0.1, 0),
                        c(0, 0.2, 0.6, 0.2, 0),
                        c(0, 0.25, 0.5, 0.25, 0))  
  }
  colnames(combos) = LETTERS[1:length(colnames(combos))]

  # Find best distribution
  cors = c()
  for (i in 1:ncol(combos)) {
    ge.subset = data.full.extras[data.full.extras$scale == target, ]
    ge.subset[ge.subset$degree == "mid", "speaker.p"] = combos[, i]
    target.scalars = c("hi1", "hi2")

    # Run model
    m1 = ge.subset %>%
      do(run.full(., alpha = alpha, addMid=T, normalize=normalize))
    if (target == "some_all") {
      target.scalars = c("hi1", "low1")
      m1 = m1[m1$degree %in% target.scalars, ]  
    } else {
      m1 = m1[m1$degree %in% target.scalars, ]  
    }
    cors[i] = cor(m1$pred, m1$listener.p)  
  }

  # Get index of highest correlation
  i = which(cors == max(cors))[1]
  ge.subset = data.full.extras[data.full.extras$scale == target, ]
    ge.subset[ge.subset$degree == "mid", "speaker.p"] = combos[, i]
  m1 = ge.subset %>%
    do(run.full(., alpha = alpha, addMid=T, normalize=normalize))

  p = qplot(stars, listener.p, col=degree, 
        data=m1,
        ylab="Posterior p(rating | word)") + 
    geom_line(aes(y = pred), lty = 4)

  return(c(combos[i], cors[i]))
}
```

```{r}
# liked_loved
combos = data.frame(c(0, 0, 1, 0, 0),
                    c(0, 0.05, 1, .05, 0),
                    c(0, 0.1, 1, .1, 0),
                    c(0, 0.1, 0.9, .1, 0),
                    c(0, 0.2, 0.8, .2, 0),
                    c(0, 0.25, 0.8, .25, 0),
                    c(0.05, 0.1, 1, .1, 0.9),
                    c(0.1, 0.1, 1, .1, 0.1))  
focus_alts("liked_loved", combos=combos, alpha=4, normalize=T)
focus_alts("liked_loved", combos=combos, alpha=8, normalize=F)

# good_excellent
combos = data.frame(c(0, 0, 1, 0, 0),
                    c(0, 0.05, 0.9, 0.05, 0),
                    c(0, 0.1, 0.8, 0.1, 0),
                    c(0, 0.2, 0.2, 0.2, 0.9),
                    c(0, 0.2, 0.6, 0.2, 1),
                    c(0, 0.2, 0.6, 0.3, 0.2),
                    c(0, 0.2, 0.6, 0.3, 0.3),
                    c(0, 0.2, 0.6, 0.5, 0.4),
                    c(0, 0.2, 0.6, 0.5, 0.5),
                    c(0, 0.2, 0.6, 0.6, 0.4),
                    c(0.05, 0.2, 0.6, 0.2, 0.05),
                    c(0, 0.25, 0.5, 0.25, 0))  
focus_alts("good_excellent", combos=combos, alpha = 1, normalize=T)
focus_alts("good_excellent", combos=combos, alpha = 8, normalize=F)

# some_all
combos = data.frame(c(0, 1, 0, 0, 0),
                    c(0.05, 2, 0.05, 0.05, 0.05),
                    c(0.2, 0.9, 0.2, 0.1, 0.1),
                    c(0.2, 0.9, 0.2, 0.1, 0),
                    c(0.1, 0.9, 0.1, 0, 0),
                    c(0.1, 0.8, 0.1, 0, 0),
                    c(0.2, 0.6, 0.2, 0, 0),
                    c(0.25, 0.5, 0.25, 0, 0))
focus_alts("some_all", combos=combos, alpha=4, normalize=T)
focus_alts("some_all", combos=combos, alpha=8, normalize=F)

# memorable_unforgettable
combos = data.frame(rep(0.1, 5),
                    c(0.05, 2, 0.05, 0.05, 0.05),
                    c(0.05, 1, 0.05, 0.05, 0.05),
                    c(0.05, 0.9, 0.05, 0.05, 0.05),
                    c(0.05, 0.8, 0.05, 0.05, 0.05),
                    c(0.05, 0.7, 0.05, 0.05, 0.05),
                    c(0.05, 0.6, 0.05, 0.05, 0.05),
                    c(0.05, 0.5, 0.05, 0.05, 0.05),
                    c(0.1, 0.8, 0.1, 0.1, 0.1))
focus_alts("memorable_unforgettable", combos=combos, alpha=4, normalize=T)
focus_alts("memorable_unforgettable", combos=combos, alpha=8, normalize=F)

# palatable_delicious
combos = data.frame(rep(0.1, 5),
                    c(0.05, 2, 0.05, 0.05, 0.05),
                    c(0.05, 1, 0.05, 0.05, 0.05),
                    c(0.05, 0.05, 0.05, 0.9, 0.1),
                    c(0.05, 0.05, 0.05, 0.9, 0.05),
                    c(0.05, 0.7, 0.05, 0.5, 0.05),
                    c(0.05, 0.6, 0.05, 0.05, 0.05),
                    c(0.05, 0.5, 0.05, 0.05, 0.05),
                    c(0.1, 0.8, 0.1, 0.1, 0.1))
focus_alts("palatable_delicious", combos=combos, alpha=4, normalize=T)
focus_alts("palatable_delicious", combos=combos, alpha=8, normalize=F)
```

>>>>

# Fit extra scalar literal listener distributions
## modified gradient descent / MLE fn()
```{r}
# normalize()
# -----------
# normalize a generic vector
normalize = function(w) { return (w / sum(w))}

# fit_distr()
# -----------
# target        --> specific scalar family we're interested in (i.e. "some_all")
# numIters      --> number of iterations to run
# eta           --> step size
# alpha         --> alpha level for RSA?
# normalize     --> normalize in RSA?
fit_distr = function(target, numIters = 100, eta=0.01, normalize=T,
                     alpha=1, verbose=F, data.record=NULL) {

  if (verbose) { print(paste("Currently fitting", target, "...")) }
  # Initialize to uniform
  w = rep(0.1, 5)
  # Focus on scalar 'target'
  scalar.focused = data.full.extras[data.full.extras$scale == target, ]

  # For numIters iterations...
  for (i in 1:numIters) {
    prev.w = w
    scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = w
    m = scalar.focused %>%
      do(run.full(., alpha = alpha, addMid=T, normalize=normalize))
    w[1] = w[1] - eta * (sum(m[m$stars == 1, ]$listener.p - m[m$stars == 1, ]$pred))
    w[2] = w[2] - eta * (sum(m[m$stars == 2, ]$listener.p - m[m$stars == 2, ]$pred))
    w[3] = w[3] - eta * (sum(m[m$stars == 3, ]$listener.p - m[m$stars == 3, ]$pred))
    w[4] = w[4] - eta * (sum(m[m$stars == 4, ]$listener.p - m[m$stars == 4, ]$pred))
    w[5] = w[5] - eta * (sum(m[m$stars == 5, ]$listener.p - m[m$stars == 5, ]$pred))
    w = ifelse(w <= 0, 0, w)
    w = ifelse(w >= 1, 1, w)
    w = normalize(w)

    # Record convergence
    if (!is.null(data.record)) {
      data.record = rbind(data.record, w)
    }

    # Check if we've converged
    if (sum(abs(prev.w - w)) < 0.00001) {
      if(verbose) { print("Convergence") }
      return(list(w, data.record))
    }

    if(verbose) {print(list(i, w))}
  }
  return(list(w, data.record))
}
```

# Fit extra scalar literal listener distributions
```{r}
# Normalize in RSA?
norm = T
# Our scalars
scalars = c("good_excellent",
            "liked_loved",
            "palatable_delicious",
            "memorable_unforgettable",
            "some_all")
fitted.distr = data.frame(matrix(rep(NA, 25), ncol=5))
colnames(fitted.distr) = scalars

# Fit each of our scalars / track fitting
track.data = data.frame(matrix(ncol=6))
for (i in 1:length(scalars)) {
  # record fitting
  data.record = data.frame(matrix(rep(NA, 5), ncol=5))
  # output is optimal distr [[1]] and [[2]] record of fitting
  out = fit_distr(scalars[i], numIters=20, eta=0.01,
                  verbose=F, normalize=norm, alpha=4, data.record=data.record)

  # get fitted scalar distr
  fitted.distr[, i] = out[[1]]
  # get tracking data
  tracking = as.data.frame(out[[2]][-1,]) %>%
    mutate(X6 = scalars[i])
  track.data = rbind(track.data, tracking)
}
```

```{r}
# Format track data for display
track.data = track.data[-1, ]
track.data$step = rep(seq(1, 20), 5)
track.data = track.data %>% 
  gather(stars, value, X1:X5)
```

## Run distribution fitting
```{r}
for (scale in scalars) {
  data.full.extras[data.full.extras$scale == scale &
                     data.full.extras$degree == "mid", ]$speaker.p =
    normalize(fitted.distr[ , scale])
}
m = data.full.extras %>%
  do(run.full(., alpha = 4, addMid=T, normalize=norm))
cor(m[matched.items.extras, ]$listener.p, m[matched.items.extras, ]$pred)
```

## Plot distr
```{r}
# Normalize speaker.p for plotting
for (scale in scalars) {
  degrees = unique(data.full.extras$degree)
  for (degree in degrees) {
    data.full.extras[data.full.extras$scale == scale &
                       data.full.extras$degree == degree, ]$speaker.p = 
        normalize(data.full.extras[data.full.extras$scale == scale &
                                     data.full.extras$degree == degree, ]$speaker.p)
  }
}

# Differentiate tuned 'mid' scalar
data.full.extras$type = ifelse(data.full.extras$degree == "mid", "red", "blue")
qplot(x = stars, y = speaker.p, colour = type,
      geom = "line", stat = "identity", position = "dodge", 
      main = "Literal listener semantics\nEmpirical estimates and tuned extra 'mid'",
      ylab = "P_{L0}(Word | Rating) - Normalized",
      data = data.full.extras) + 
  facet_wrap(~scale + degree)

# Tracking of gradient descent tuning
ggplot(data=track.data, aes(x=step, y=value, col=stars)) +
  geom_line() +
  facet_wrap(~X6)
```
